---
title: "analyze_rnaseq"
author: "Natalie Davidson"
date: "3/11/2021"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

require("data.table")
require("here")
require("ggplot2")
require("dplyr")
require("readr")
require("SummarizedExperiment")
require("ggrepel")
require("ggcorrplot")
require("biomaRt")
require("VennDiagram")
require("pheatmap")
require("RColorBrewer")

proj_dir = here()
source(file.path(proj_dir, "/analyze_rnaseq/plot_utils.R"))
source(file.path(proj_dir, "/comparison/utils/file_processing_utils.R"))
set.seed(5)

```
## RNASeq after Salmon Quant QC

This document is for QC checking for any anomalies in the RNA-Seq counts:

Subtypes described in here are from the file: `/reference_data/RNAseq_sample_selection_subtable_removed.tsv`
This file contains the NanoString predicted subtypes.
This file was received in the email from Mollie Barnard "RNAseq pilot sample summary"
The original file is called "RNAseq sample selection.csv" and has the additional
slideshow "NCOCS RNAseq pilot sample summary".


### read in the RDS file containing gene counts


```{r qc_matrs, fig.align='center', echo=TRUE, warning=FALSE, message=F}

#306 samples in the full study were sent for sequencing.

#The following samples failed: 

samps_failed = c("Sample_19291X9", "Sample_19291X35", "Sample_19291X57", "Sample_19291X124", "Sample_19291X175")
failed_df = data.frame(metric=rep("failed", length(samps_failed)),
                       samp_ids = samps_failed)


#The following samples were re-sequenced to get higher yeild:
samps_reseq = c("Sample_19291X8", "Sample_19291X20", "Sample_19291X36", "Sample_19291X37", 
                 "Sample_19291X43", "Sample_19291X60", "Sample_19291X165", "Sample_19291X187",
                "Sample_19291X188", "Sample_19291X205", "Sample_19291X206", "Sample_19291X258", 
                "Sample_19291X262", "Sample_19291X263")
re_seq_df = data.frame(metric=rep("re_seq", length(samps_reseq)),
                       samp_ids = samps_reseq)


#The following samples got fewer than 35M reads
samps_35M = c("Sample_19291X20", "Sample_19291X26", "Sample_19291X36", "Sample_19291X40", 
              "Sample_19291X73", "Sample_19291X100", "Sample_19291X109", "Sample_19291X165", 
              "Sample_19291X297", "Sample_19291X303")
u35M_df = data.frame(metric=rep("u35M", length(samps_35M)),
                       samp_ids = samps_35M)

# the following had low starting material
samps_low_start = c("Sample_19291X38", "Sample_19291X39", "Sample_19291X40", 
                    "Sample_19291X43", "Sample_19291X57", "Sample_19291X60", 
                    "Sample_19291X100", "Sample_19291X102", "Sample_19291X108", 
                    "Sample_19291X124", "Sample_19291X164", "Sample_19291X175", 
                    "Sample_19291X178", "Sample_19291X190", "Sample_19291X197", 
                    "Sample_19291X206", "Sample_19291X227", "Sample_19291X236", 
                    "Sample_19291X237", "Sample_19291X247", "Sample_19291X250", 
                    "Sample_19291X252", "Sample_19291X276", "Sample_19291X293", 
                    "Sample_19291X297")
low_start_df = data.frame(metric=rep("low_start", length(samps_low_start)),
                       samp_ids = samps_low_start)


qc_degradation = do.call(rbind, list(failed_df, re_seq_df, u35M_df, low_start_df))

# add visualization
display_venn(
  list(samps_failed, samps_reseq, samps_35M, samps_low_start),
  category.names = c("Failed" , "Re-Sequenced", "under35M", "low starting material"),
  fill = brewer.pal(n = 4, name = "Set3")
  )


```



### read in the RDS file containing gene counts


```{r read_in, fig.align='center', echo=FALSE, warning=FALSE, message=F}

get_count <- function(infile, get_dates=T){

    gene_summarized = readr::read_rds(infile)
    count_matr = assay(gene_summarized, "counts")
    
    df_dates=NA
    if(get_dates){
        df_dates = gene_summarized$date
    }
    return(list(count_matr, df_dates))
}


# get the project directory, everything is set relative to this
proj_dir = here()

# read in original RNASeq files
gene_expr_file = file.path(proj_dir, 
                    "/data/rna_seq/salmon_quant_processed/salmon_gene_quant.RDS")

gene_count_orig = get_count(gene_expr_file)[[1]]

# read in new RNASeq files
gene_count_new = NA
date_new = NA
for(idx in 1:9){
    curr_file = paste0("salmon_gene_quant_", idx, ".RDS")
    gene_expr_file = file.path(proj_dir, 
                    "/data/rna_seq_full/salmon_quant_processed/", curr_file)
    res = get_count(gene_expr_file)
    new_df = res[[1]]
    date_df = res[[2]]

    if(!is.na(gene_count_new)){
        gene_count_new = cbind(gene_count_new, new_df)
        date_new = rbind(date_new, data.frame(date=date_df))
    }else{
        gene_count_new = new_df
        date_new = data.frame(date=date_df)
    }

}

```


# now we need to process the samples that were resequenced

The samples that were re-run are all on lane 3, the original sequencing of them
were done on either lane 1, 2, or 3.
To disambiguate them we select on the most recently sequenced sample


```{r remove_reseq, fig.align='center', echo=TRUE, warning=FALSE, message=F}

dim(gene_count_new)

idx_remove = c()
for(curr_samp_id in samps_reseq){
    
    # get the duplicate columns
    dup_idx = grep(curr_samp_id, colnames(gene_count_new))

    # chose the one sequenced later
    # assumes only 2 runs
    idx_min = which.min(as.numeric(date_new[dup_idx,]))
    
    idx_remove = c(idx_remove, dup_idx[idx_min])
}

# now remove them
gene_count_new = gene_count_new[,-idx_remove]
dim(gene_count_new)

assertthat::are_equal(301, ncol(gene_count_new))

# remove the lane info
colnames(gene_count_new) = substr(colnames(gene_count_new), 1, nchar(colnames(gene_count_new)) - 5)

# write out the file
gene_count_se = SummarizedExperiment(assays=list(counts=gene_count_new),
                                        rowData=data.frame(gene_id=row.names(gene_count_new)), 
                                        colData=data.frame(gnomex_id = colnames(gene_count_new)))
gene_expr_file = file.path(proj_dir, 
                    "/data/rna_seq_full/salmon_quant_processed/salmon_gene_quant.RDS")
saveRDS(gene_count_se, file = gene_expr_file)

```

Merge together the new and old data sources

```{r combine_new_old, fig.align='center', echo=FALSE, warning=FALSE, message=F}

# now join them together
# make sure rownames are the same
stopifnot(all(row.names(gene_count_new) == row.names(gene_count_orig)))
gene_count = cbind(gene_count_orig, gene_count_new)


# read in the samples that should be excluded
# we will remove all samples that are from white women
samp_exclusion_file = file.path(proj_dir, 
                    "/reference_data/way_exclusions.csv")
samp_exclusion_df = data.frame(fread(samp_exclusion_file))
samp_exclusion_df = subset(samp_exclusion_df, exclude_white==TRUE)

# remove the samples that are from white women
sample_ids_remove = paste0("Sample_", samp_exclusion_df$ID)
sample_ids_keep = setdiff(colnames(gene_count), sample_ids_remove)
stopifnot(length(sample_ids_keep) == 325)
gene_count = gene_count[,sample_ids_keep]

# remove the samples that have neoadjuvant therapy
neoadj_file = file.path(proj_dir, 
                    "/reference_data/tissuegrant_epidata_08102022.csv")
neoadj_df = data.frame(fread(neoadj_file))
neoadj_df = subset(neoadj_df, neoadj=="yes")

rnaseq_trans_file = file.path(proj_dir, 
                    "/reference_data/AACES_FFPE_RNA_JeffMarks_Doherty.tsv")
rnaseq_trans_df = data.frame(fread(rnaseq_trans_file))
neoadj_df = merge(neoadj_df[,c("suid", "neoadj")], rnaseq_trans_df[,c("suid", "ID")])
assertthat::are_equal(nrow(neoadj_df), 53)

sample_ids_remove = paste0("Sample_", neoadj_df$ID)
sample_ids_keep = setdiff(colnames(gene_count), sample_ids_remove)
stopifnot(length(sample_ids_keep) == 272)
gene_count = gene_count[,sample_ids_keep]

# add in a tag for which experiment the sample is from
expr_version = c(rep("orig", length(intersect(colnames(gene_count_orig), sample_ids_keep))),
                 rep("new", length(intersect(colnames(gene_count_new), sample_ids_keep))))
expr_version_df = data.frame(version=expr_version, sample_id=colnames(gene_count))

# add in other metadata
expr_version_df$re_seq = FALSE
expr_version_df$re_seq[expr_version_df$sample_id %in% samps_reseq] = TRUE

expr_version_df$u35M = FALSE
expr_version_df$u35M[expr_version_df$sample_id %in% samps_35M] = TRUE

expr_version_df$low_start = FALSE
expr_version_df$low_start[expr_version_df$sample_id %in% samps_low_start] = TRUE


gene_count[1:5,1:5]

```

### plot general stats about the samples

Here we want to see if there are any sample outliers in expression that may be due to technical artifacts.
So lets look at the raw count values and see if any samples have much higher expression than others.
```{r raw_count_comparison, fig.align='center', fig.width=15, fig.height=7, echo=FALSE, warning=FALSE, message=F}

# now plot the stats
gene_count_df = data.frame(gene_count)
gene_count_df$gene_id = row.names(gene_count_df)
gene_count_melt <- melt(gene_count_df, id.vars="gene_id")
colnames(gene_count_melt) = c("gene_id", "sample_id", "count")

# add version tag
gene_count_melt = merge(gene_count_melt, expr_version_df)

# order samples by median expression
sample_order_df = data.frame(median_count = apply(gene_count, 2, median),
                             sample_id = colnames(gene_count))
sample_order_df = sample_order_df[order(sample_order_df$median_count),]
gene_count_melt$sample_id = factor(gene_count_melt$sample_id, 
                                    levels = sample_order_df$sample_id)
ggplot(gene_count_melt, aes(x=as.factor(sample_id), y=count, fill=version)) +
    geom_boxplot() + theme_bw() +
    theme(axis.text.x = element_blank()) +
    xlab("Sample IDs") + ylab("count") +
        ggtitle("count distribution over gene for each sample")

ggplot(gene_count_melt, aes(x=as.factor(sample_id), y=log10(count+1), fill=version)) +
    geom_boxplot() + theme_bw() +
    theme(axis.text.x = element_blank()) +
    xlab("Sample IDs") + ylab("log10(count+1)") +
        ggtitle("count distribution over gene for each sample")


```

It's a bit difficult to see, to lets filter out genes that have too low expression.
To determine this, lets look at the number of genes remaining at different median cutoffs

```{r gene_filtering_by_expr, fig.align='center', width=15, height=7, echo=FALSE, warning=FALSE, message=F}

# now plot the stats
cutoff_df = NA
for(med_cutoff in 0:10){
    
    filtered_genes = length(which(apply(gene_count, 1, median)<=med_cutoff))
    cutoff_res = c(filtered_genes/nrow(gene_count), 
                   1-filtered_genes/nrow(gene_count),
                   med_cutoff)
    cutoff_df = rbind(cutoff_df, cutoff_res)

}
cutoff_df = data.frame(cutoff_df[-1,])
colnames(cutoff_df) = c("genes_removed", "genes_kept", "cutoff")
cutoff_df$cutoff = factor(cutoff_df$cutoff)
cutoff_df = melt(cutoff_df, id.vars=c("cutoff"))

ggplot(cutoff_df, aes(x=cutoff, y=value, fill=variable)) +
   geom_bar(stat="identity", position="stack") + theme_bw() +
    theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1)) +
    ylab("% Genes") + xlab("cutoff") +
        ggtitle("% of genes remaining at different cutoffs")

```

A cutoff of greater than zero seems to maintain a more than 50% of genes, so lets now replot with this filtering.
Additionally, using this level of filtering lets normalize by 85th quantile.
We see that samples 43, 42, 35 have the lowest expression overall.
We also see that the samples with low starting material are also in the bottom,
so we filter out the bottom 10 samples.

```{r filt_normalized_count_comparison, fig.align='center', fig.width=15, fig.height=7, echo=FALSE, warning=FALSE, message=F}

# now plot the filtered version
filtered_genes = rownames(gene_count)[which(apply(gene_count, 1, median)==0)]
gene_count_df_filt = subset(gene_count_melt, ! gene_id %in% filtered_genes)

ggplot(gene_count_df_filt, aes(x=as.factor(sample_id), y=log10(count+1), fill=version)) +
    geom_boxplot() + theme_bw() +
    theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1)) +
    xlab("Sample IDs") + ylab("log10(count+1)") +
        ggtitle("filtered count distribution over gene for each sample")

# now plot the normalized version
gene_count_filt = subset(gene_count, 
                         ! rownames(gene_count) %in% filtered_genes)
samp_85th = apply(gene_count_filt, 2, quantile, 0.85)
samp_85th = samp_85th/max(samp_85th)
gene_count_filt_norm = sweep(gene_count_filt, 2, samp_85th, FUN = '/')

gene_count_filt_norm_melt = data.frame(gene_count_filt_norm)
gene_count_filt_norm_melt$gene_id = row.names(gene_count_filt_norm_melt)
gene_count_filt_norm_melt <- melt(gene_count_filt_norm_melt, id.vars="gene_id")
colnames(gene_count_filt_norm_melt) = c("gene_id", "sample_id", "norm_count")

# add version tag
gene_count_filt_norm_melt = merge(gene_count_filt_norm_melt, expr_version_df)

# order samples by median expression
sample_order_df = data.frame(median_count = apply(gene_count_filt_norm, 2, median),
                             sample_id = colnames(gene_count_filt_norm))
sample_order_df = sample_order_df[order(sample_order_df$median_count),]

# save the samples to remove
SAMPLES_FILTERED_OUT = sample_order_df$sample_id[1:10]
gene_count_filt_norm_melt$sample_id = factor(gene_count_filt_norm_melt$sample_id, 
                                    levels = sample_order_df$sample_id)



ggplot(gene_count_filt_norm_melt, aes(x=as.factor(sample_id), y=log10(norm_count+1), fill=version)) +
    geom_boxplot() + theme_bw() +
    theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1)) +
    xlab("Sample IDs") + ylab("log10(norm_count+1)") +
        ggtitle("filtered + normalized count distribution over gene for each sample")

ggplot(gene_count_filt_norm_melt, aes(x=as.factor(sample_id), y=log10(norm_count+1), fill=re_seq)) +
    geom_boxplot() + theme_bw() +
    theme(axis.text.x = element_blank()) +
    xlab("Sample IDs") + ylab("log10(count+1)") +
        ggtitle("count distribution over gene for each sample")

ggplot(gene_count_filt_norm_melt, aes(x=as.factor(sample_id), y=log10(norm_count+1), fill=u35M)) +
    geom_boxplot() + theme_bw() +
    theme(axis.text.x = element_blank()) +
    xlab("Sample IDs") + ylab("log10(count+1)") +
        ggtitle("count distribution over gene for each sample")

ggplot(gene_count_filt_norm_melt, aes(x=as.factor(sample_id), y=log10(norm_count+1), fill=low_start)) +
    geom_boxplot() + theme_bw() +
    theme(axis.text.x = element_blank()) +
    xlab("Sample IDs") + ylab("log10(count+1)") +
        ggtitle("count distribution over gene for each sample")
```

Write out the data with the appropriate annotations

```{r write_out_processed_data, echo=TRUE, warning=FALSE, message=F}


samp_metadata = expr_version_df
row.names(samp_metadata) = samp_metadata$sample_id
samp_metadata = samp_metadata[colnames(gene_count),]

# write out the table with sample level annotations
gene_count_se = SummarizedExperiment(assays=list(counts=gene_count),
                                        rowData=data.frame(gene_id=row.names(gene_count_new)), 
                                        colData=samp_metadata)
gene_expr_file = file.path(proj_dir, 
                    "/data/rna_seq_pilot_and_new/salmon_gene_quant_full_annotated.RDS")
saveRDS(gene_count_se, file = gene_expr_file)



```

Now lets do a PCA of ALL samples.

```{r pca_filt_norm, fig.align='center', fig.width=10, fig.height=10, echo=FALSE, warning=FALSE, message=F}

color_ver_df = expr_version_df
colnames(color_ver_df)[2] = "Sample"
plot_pca(gene_count_filt_norm, title = "Filtered+Normalized PCA",
         color_df = color_ver_df, color_id = "version")

plot_pca(gene_count_filt_norm, title = "Filtered+Normalized PCA",
         color_df = color_ver_df, color_id = "re_seq")
plot_pca(gene_count_filt_norm, title = "Filtered+Normalized PCA",
         color_df = color_ver_df, color_id = "u35M")
plot_pca(gene_count_filt_norm, title = "Filtered+Normalized PCA",
         color_df = color_ver_df, color_id = "low_start")

```

Now lets make a correlation map across samples.

```{r corr_map, fig.align='center', fig.width=15, fig.height=15, echo=FALSE, warning=FALSE, message=F}


rnaseq_expr_file = file.path(proj_dir, 
                    "/data/rna_seq/salmon_quant_processed/salmon_gene_quant.RDS")

rnaseq_trans_file = file.path(proj_dir, 
                    "/reference_data/rna_sample_metadata.txt")
rnaseq_dt_orig = format_rnaseq_data(rnaseq_expr_file, rnaseq_trans_file)


rnaseq_expr_file = file.path(proj_dir, 
                    "/data/rna_seq_full/salmon_quant_processed/salmon_gene_quant.RDS")
rnaseq_trans_file = file.path(proj_dir, 
                    "/reference_data/AACES_FFPE_RNA_JeffMarks_Doherty.tsv")
extra_meta_df = expr_version_df
extra_meta_df$ID = gsub("Sample_", "", extra_meta_df$sample_id)
extra_meta_df = extra_meta_df[,c("version", "ID", "re_seq", "u35M", "low_start")]
rnaseq_dt_new = format_rnaseq_data(rnaseq_expr_file, 
                                   rnaseq_trans_file,
                                   extra_meta_df)
rnaseq_dt_orig$version = "pilot"
rnaseq_dt_orig$re_seq = FALSE
rnaseq_dt_orig$u35M = FALSE
rnaseq_dt_orig$low_start = FALSE

rnaseq_dt = rbind(rnaseq_dt_orig, rnaseq_dt_new)
rnaseq_dt$suid = as.character(rnaseq_dt$suid)

# sample subtype annotation
subtype_file = file.path(proj_dir, 
                    "/reference_data/RNAseq_sample_selection_subtable_removed.tsv")
subtype_dt = fread(subtype_file)
subtype_dt = unique(subtype_dt[,c("suid", "HGSCsubtype")])
subtype_dt$suid = as.character(subtype_dt$suid)

total_dt = merge(subtype_dt, rnaseq_dt, all.y=T, by="suid")

total_dt$HGSCsubtype[which(is.na(total_dt$HGSCsubtype))] = "unknown"



# remove the samples that are from white women
sample_ids_remove = samp_exclusion_df$ID
sample_ids_keep = setdiff(unique(total_dt$ID), sample_ids_remove)
stopifnot(length(sample_ids_keep) == 325)
total_dt = subset(total_dt, ID %in% sample_ids_keep)

# remove neoadjuvant samples
sample_ids_remove = neoadj_df$ID
sample_ids_keep = setdiff(unique(total_dt$ID), sample_ids_remove)
stopifnot(length(sample_ids_keep) == 272)
total_dt = subset(total_dt, ID %in% sample_ids_keep)


samp_meta_file = file.path(proj_dir, 
                    "/data/rna_seq_pilot_and_new/sample_metadata.tsv")
samp_meta_df = subset(total_dt, select=-c(ensembl_gene_id, hgnc_symbol, expr))
samp_meta_df = unique(samp_meta_df)
write.table(samp_meta_df, samp_meta_file, quote=F, sep="\t", row.names=F)

# to make comparable scale within each sample
total_dt_scaled = total_dt %>%
                        group_by(suid) %>%
                        mutate(scaled_expr = scale(expr))

# now dcast so we can analyze them
total_dt_cast = dcast(total_dt_scaled, suid+ID+HGSCsubtype+version+re_seq+u35M+low_start ~ ensembl_gene_id, 
                      value.var="scaled_expr", fun.aggregate = mean)
row.names(total_dt_cast) = total_dt_cast$suid

# now plot heatmap across all genes
expr_df = subset(total_dt_cast, select=-c(suid, ID, HGSCsubtype, version, re_seq, u35M, low_start))

annotation_row_df <- data.frame(subset(total_dt_cast, select=c(HGSCsubtype, version, re_seq, u35M, low_start)))
annotation_row_df$HGSCsubtype = as.factor(annotation_row_df$HGSCsubtype)
annotation_row_df$version = as.factor(annotation_row_df$version)
annotation_row_df$re_seq = as.factor(annotation_row_df$re_seq)
annotation_row_df$u35M = as.factor(annotation_row_df$u35M)
annotation_row_df$low_start = as.factor(annotation_row_df$low_start)


# now plot heatmap of correlation matrix
cor_df <- cor(t(expr_df), method="spearman", use="complete.obs")
pheatmap(cor_df, annotation_row = annotation_row_df, 
         annotation_col = annotation_row_df, 
         scale = "none", show_rownames = T,
         show_colnames = F, cluster_cols = T, cluster_rows = T,
         main = "Filtered+Normalized spearman corr. matr. all samples")



```


Now lets restrict the analysis to only look at the expression of the genes that are needed in the UBC classifier.
First lets see how many of these genes have median expression > 0

All genes pass.

```{r ubc_filtering, fig.align='center', echo=FALSE, warning=FALSE, message=F}

# get the UBC file
ubc_file = file.path(proj_dir, "/reference_data/UBC_genes.txt")
ubc_df <- data.frame(fread(ubc_file))

# translate the gene names
colnames(ubc_df)[1] = "hgnc_symbol"
ubc_gene_ids_df = get_gene_id_map(unique(ubc_df$hgnc_symbol), 
                                filter_type="hgnc_symbol",
                                attributes= c("hgnc_symbol", "ensembl_gene_id"))


gene_count_filt_norm_UBC = subset(total_dt, 
                                  total_dt$ensembl_gene_id %in% 
                                      ubc_gene_ids_df$ensembl_gene_id)

# show the venn diagram
display_venn(
  list(ubc_gene_ids_df$ensembl_gene_id, rownames(gene_count_filt_norm)),
  category.names = c("UBC" , "Filtered Genes"),
  fill = c("#999999", "#009E73")
  )

print("Missing Genes")
subset(ubc_gene_ids_df, 
          !ubc_gene_ids_df$ensembl_gene_id %in% 
           rownames(gene_count_filt_norm))

duplicated_genes = ubc_gene_ids_df$hgnc_symbol[duplicated(ubc_gene_ids_df$hgnc_symbol)]
print("Duplicated Genes due to ENSG translation")

subset(ubc_gene_ids_df, 
          ubc_gene_ids_df$hgnc_symbol %in% 
           duplicated_genes)


```

Now re-do the analysis on this constricted set of genes and all samples

```{r corr_map_ubc, fig.align='center', fig.width=10, fig.height=10, echo=FALSE, warning=FALSE, message=F}

# now dcast so we can analyze them
ubc_dt_cast = dcast(gene_count_filt_norm_UBC, suid+HGSCsubtype+version+re_seq+u35M+low_start ~ ensembl_gene_id, value.var="expr")
row.names(ubc_dt_cast) = ubc_dt_cast$suid

# now plot heatmap across all genes
ubc_expr_df = subset(ubc_dt_cast, select=-c(suid, HGSCsubtype, version, re_seq, u35M, low_start))
row.names(ubc_expr_df) = ubc_dt_cast$suid

annotation_row_df <- data.frame(subset(ubc_dt_cast, select=c(HGSCsubtype, version, re_seq, u35M, low_start)))
annotation_row_df$HGSCsubtype = as.factor(annotation_row_df$HGSCsubtype)
annotation_row_df$version = as.factor(annotation_row_df$version)
annotation_row_df$re_seq = as.factor(annotation_row_df$re_seq)
annotation_row_df$u35M = as.factor(annotation_row_df$u35M)
annotation_row_df$low_start = as.factor(annotation_row_df$low_start)
row.names(annotation_row_df) = row.names(ubc_dt_cast)

# now plot heatmap of correlation matrix
cor_df <- data.frame(cor(t(ubc_expr_df), method="spearman"))
row.names(cor_df) = row.names(ubc_expr_df)

pheatmap(cor_df, annotation_row = annotation_row_df, 
         scale = "none", show_rownames = T,
         show_colnames = F, cluster_cols = T, cluster_rows = T,
         main = "Filtered+Normalized spearman corr. matr.")


# now plot pca
pca_ubc_df = t(ubc_expr_df)
colnames(pca_ubc_df) = paste0("Sample_", ubc_dt_cast$suid)
color_df = subset(ubc_dt_cast, select=c(suid, HGSCsubtype, version, re_seq, u35M, low_start))
color_df$HGSCsubtype = as.factor(color_df$HGSCsubtype)
color_df$version = as.factor(color_df$version)
color_df$re_seq = as.factor(color_df$re_seq)
color_df$u35M = as.factor(color_df$u35M)
color_df$low_start = as.factor(color_df$low_start)

color_df$suid = paste0("Sample_", color_df$suid)
colnames(color_df)[1] = "Sample"

plot_pca(pca_ubc_df, title = "PCA of normalized genes, only UBC genes",
         color_df = color_df, color_id = "HGSCsubtype")

plot_pca(pca_ubc_df, title = "PCA of normalized genes, only UBC genes",
         color_df = color_df, color_id = "version")

plot_pca(pca_ubc_df, title = "PCA of normalized genes, only UBC genes",
         color_df = color_df, color_id = "re_seq")

plot_pca(pca_ubc_df, title = "PCA of normalized genes, only UBC genes",
         color_df = color_df, color_id = "u35M")

plot_pca(pca_ubc_df, title = "PCA of normalized genes, only UBC genes",
         color_df = color_df, color_id = "low_start")

plot_pca(pca_ubc_df, title = "PCA of normalized genes, only UBC genes")



```


