---
title: "compare_clusters"
author: "Natalie Davidson"
date: "8/10/2021"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

require("data.table")
require("here")
require("ggplot2")
require("dplyr")
require("caret")
library(ggpubr)
theme_set(theme_pubr())

proj_dir = here()

set.seed(5)

```

# Intro

In this notebook I compare the clusters from the published output of the Way pipeline and my run.
Since the clusters are mapped to TCGA clusters, we can use the cluster IDs to compare them.
To do this, I treated the cluster IDs as expected (published) and predicted (my run) labels.
This enables me to calculate balanced accuracy between the published results and my pipeline results.
We do this across all datasets and in Kmeans and NMF.

This notebook contains a lot of detailed results, but the important results are at the bottom in the section "Balanced Accuracy Summary".
The summary contains the balanced accuracy across all datasets and sizes of K (2-4).
Balanced accuracy is used when the cluster sizes differ -- which in this case they do slightly.

It is calculated as (Sensitivity + Specificity) / 2

Sensitivity (true positive rate) = TP / (TP + FN) 

Specificity (true negative rate) = TN / (TN + FT) 

# First read the files and process

```{r read_in_files, fig.align='center', fig.width=15, fig.height=15, echo=TRUE, warning=FALSE, message=F}

new_file = file.path(proj_dir, 
                    "/data/way_pipeline_results_10removed_NeoRemoved/2.Clustering_DiffExprs-Tables-ClusterMembership/FullClusterMembership.csv")

new_df = data.frame(fread(new_file))
colnames(new_df)[1] = "ID"

old_file = file.path(proj_dir, 
                    "/data/way_publication_results/033514_tables3.csv")

old_df = data.frame(fread(old_file))
old_df$Dataset[which(old_df$Dataset == "Mayo")] = "mayo.eset"

# only take samples that are in both
keep_ids = intersect(new_df$ID, old_df$ID)
new_df = subset(new_df, ID %in% keep_ids)
old_df = subset(old_df, ID %in% keep_ids)

# order them so we can directly compare
new_df = new_df[order(new_df$ID), ]
old_df = old_df[order(old_df$ID), ]
stopifnot(all(new_df$ID == old_df$ID))

```


# Plot confusion matrices

```{r conf_matr_method, fig.align='center', fig.width=15, fig.height=15, echo=TRUE, warning=FALSE, message=F}

get_conf_matrix <- function(curr_dataset, curr_clust, new_df, old_df){
    

    expected_val = subset(old_df, Dataset == curr_dataset)
    predicted_val = subset(new_df, Dataset == curr_dataset)
    stopifnot(all(expected_val$ID == predicted_val$ID))
    
    expected_val = expected_val[,curr_clust]
    predicted_val = predicted_val[,curr_clust]
    
    conf_res <- confusionMatrix(data=as.factor(predicted_val), 
                                reference = as.factor(expected_val))
    return(conf_res)

}


```

### TCGA Kmeans

```{r plot_tcga, fig.align='center', fig.width=15, fig.height=15, echo=TRUE, warning=FALSE, message=F}



print("K=2")
conf_res2 <- get_conf_matrix("TCGA", "ClusterK2_kmeans", new_df, old_df)
print(conf_res2)


print("K=3")
conf_res3 <- get_conf_matrix("TCGA", "ClusterK3_kmeans", new_df, old_df)
print(conf_res3)


print("K=4")
conf_res4 <- get_conf_matrix("TCGA", "ClusterK4_kmeans", new_df, old_df)
print(conf_res4)

BA_tcga_kmeans = c(conf_res2$byClass[11], conf_res3$byClass[,11], conf_res4$byClass[,11])
BA_tcga_kmeans

```

### TCGA NMF

```{r plot_tcga_nmf, fig.align='center', fig.width=15, fig.height=15, echo=TRUE, warning=FALSE, message=F}

print("K=2")
conf_res2 <- get_conf_matrix("TCGA", "ClusterK2_NMF", new_df, old_df)
print(conf_res2)


print("K=3")
conf_res3 <- get_conf_matrix("TCGA", "ClusterK3_NMF", new_df, old_df)
print(conf_res3)


print("K=4")
conf_res4 <- get_conf_matrix("TCGA", "ClusterK4_NMF", new_df, old_df)
print(conf_res4)

BA_tcga_nmf = c(conf_res2$byClass[11], conf_res3$byClass[,11], conf_res4$byClass[,11])
BA_tcga_nmf

```


### Mayo Kmeans

```{r plot_mayo_kmeans, fig.align='center', fig.width=15, fig.height=15, echo=TRUE, warning=FALSE, message=F}

print("K=2")
conf_res2 <- get_conf_matrix("mayo.eset", "ClusterK2_kmeans", new_df, old_df)
print(conf_res2)


print("K=3")
conf_res3 <- get_conf_matrix("mayo.eset", "ClusterK3_kmeans", new_df, old_df)
print(conf_res3)


print("K=4")
conf_res4 <- get_conf_matrix("mayo.eset", "ClusterK4_kmeans", new_df, old_df)
print(conf_res4)

BA_mayo_kmeans = c(conf_res2$byClass[11], conf_res3$byClass[,11], conf_res4$byClass[,11])
BA_mayo_kmeans

```

### Mayo NMF

```{r plot_mayo_nmf, fig.align='center', fig.width=15, fig.height=15, echo=TRUE, warning=FALSE, message=F}

print("K=2")
conf_res2 <- get_conf_matrix("mayo.eset", "ClusterK2_NMF", new_df, old_df)
print(conf_res2)


print("K=3")
conf_res3 <- get_conf_matrix("mayo.eset", "ClusterK3_NMF", new_df, old_df)
print(conf_res3)


print("K=4")
conf_res4 <- get_conf_matrix("mayo.eset", "ClusterK4_NMF", new_df, old_df)
print(conf_res4)

BA_mayo_nmf = c(conf_res2$byClass[11], conf_res3$byClass[,11], conf_res4$byClass[,11])
BA_mayo_nmf


```


### Tothill Kmeans

```{r plot_Tothill_kmeans, fig.align='center', fig.width=15, fig.height=15, echo=TRUE, warning=FALSE, message=F}

print("K=2")
conf_res2 <- get_conf_matrix("Tothill", "ClusterK2_kmeans", new_df, old_df)
print(conf_res2)


print("K=3")
conf_res3 <- get_conf_matrix("Tothill", "ClusterK3_kmeans", new_df, old_df)
print(conf_res3)


print("K=4")
conf_res4 <- get_conf_matrix("Tothill", "ClusterK4_kmeans", new_df, old_df)
print(conf_res4)

BA_tothill_kmeans = c(conf_res2$byClass[11], conf_res3$byClass[,11], conf_res4$byClass[,11])
BA_tothill_kmeans

```

### Tothill NMF

```{r plot_Tothill_nmf, fig.align='center', fig.width=15, fig.height=15, echo=TRUE, warning=FALSE, message=F}

print("K=2")
conf_res2 <- get_conf_matrix("Tothill", "ClusterK2_NMF", new_df, old_df)
print(conf_res2)


print("K=3")
conf_res3 <- get_conf_matrix("Tothill", "ClusterK3_NMF", new_df, old_df)
print(conf_res3)


print("K=4")
conf_res4 <- get_conf_matrix("Tothill", "ClusterK4_NMF", new_df, old_df)
print(conf_res4)

BA_tothill_nmf = c(conf_res2$byClass[11], conf_res3$byClass[,11], conf_res4$byClass[,11])
BA_tothill_nmf

```


### Yoshihara Kmeans

```{r plot_Yoshihara_kmeans, fig.align='center', fig.width=15, fig.height=15, echo=TRUE, warning=FALSE, message=F}

print("K=2")
conf_res2 <- get_conf_matrix("Yoshihara", "ClusterK2_kmeans", new_df, old_df)
print(conf_res2)


print("K=3")
conf_res3 <- get_conf_matrix("Yoshihara", "ClusterK3_kmeans", new_df, old_df)
print(conf_res3)


print("K=4")
conf_res4 <- get_conf_matrix("Yoshihara", "ClusterK4_kmeans", new_df, old_df)
print(conf_res4)


BA_yoshihara_kmeans = c(conf_res2$byClass[11], conf_res3$byClass[,11], conf_res4$byClass[,11])
BA_yoshihara_kmeans


```

### Yoshihara NMF

```{r plot_Yoshihara_nmf, fig.align='center', fig.width=15, fig.height=15, echo=TRUE, warning=FALSE, message=F}

print("K=2")
conf_res2 <- get_conf_matrix("Yoshihara", "ClusterK2_NMF", new_df, old_df)
print(conf_res2)


print("K=3")
conf_res3 <- get_conf_matrix("Yoshihara", "ClusterK3_NMF", new_df, old_df)
print(conf_res3)


print("K=4")
conf_res4 <- get_conf_matrix("Yoshihara", "ClusterK4_NMF", new_df, old_df)
print(conf_res4)


BA_yoshihara_nmf = c(conf_res2$byClass[11], conf_res3$byClass[,11], conf_res4$byClass[,11])
BA_yoshihara_nmf


```

# Balanced Accuracy Summary

```{r bs_summ, fig.align='center', fig.width=15, fig.height=15, echo=TRUE, warning=FALSE, message=F}

class_ids = c("K=2", "K=3, 1", "K=3, 2", "K=3, 3", "K=4, 1", "K=4, 2", "K=4, 3", "K=4, 4")


# Balanced Accuracy for Kmeans
ba_df = data.frame(class_ids,
                    BA_tcga_kmeans,
                    BA_mayo_kmeans,
                    BA_tothill_kmeans,
                    BA_yoshihara_kmeans)
ba_df


# Balanced Accuracy for NMF
ba_df = data.frame(class_ids,
                    BA_tcga_nmf,
                    BA_mayo_nmf,
                    BA_tothill_nmf,
                    BA_yoshihara_nmf)
ba_df



```

# Conclusion

We find that overall the clusters align between the previously published Way et. al. run and my run of the pipeline. 
NOTE: There exist slight data and possibly code differences between our two runs, so we do not expect 100% matching between runs.
However, we do expect high coherence between runs.
We find that the clusters are more consistent in K-Means than NMF.
We also find that K=4 is the most inconsistent between the two runs of the pipeline.
We will consider K=4 and KMeans results for remaining main analyses.
