---
title: "K3_kmeans_vs_nmf"
author: "Natalie Davidson"
date: "9/21/2022"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

require("data.table")
require("here")
require("ggplot2")
require("dplyr")
require("ggpubr")

#library("magick")
#library("pdftools")

theme_set(theme_pubr())

proj_dir = here()

set.seed(5)

```

# Recreate Figure 2 -- NMF vs KMeans

Figure 1 shows consistency of clustering approach across datasets.
So here we want to compare consistency between methods for EACH dataset.
We will be comparing NMF vs Kmeans.
We will only use K=3 for easier visualization for the main figures.


## First get file paths

```{r read_in_nmf_files, fig.align='center', fig.width=15, fig.height=15, echo=TRUE, warning=FALSE, message=F}

fig2_kmeans_dir = file.path(proj_dir, 
                    "/data/way_pipeline_results_10removed_NeoRemoved_inclWhites/2.Clustering_DiffExprs-Tables-DScores/")
fig2_kmeans_files = list.files(fig2_kmeans_dir,full.names = T )

fig2_nmf_dir = file.path(proj_dir, 
                    "/data/way_pipeline_results_10removed_NeoRemoved_inclWhites/2.Clustering_DiffExprs-Figures-nmf-DscoreVectors/")
fig2_nmf_files = list.files(fig2_nmf_dir,full.names = T )


```


## Plotting method + correlation calculation

```{r plot_func_nmf_v_kmeans, fig.align='center', fig.width=15, fig.height=15, echo=TRUE, warning=FALSE, message=F}

plot_nmf_v_kmeans_internal <- function(cor_df, curr_k){
        

    # get the comparisons of interest
    col_idx = grep(curr_k, colnames(cor_df))
    row_idx = grep(curr_k, rownames(cor_df))
    
    curr_df = cor_df[row_idx, col_idx]
    colnames(curr_df) = 1:ncol(curr_df)
    curr_df$compare_samp = row.names(curr_df) 

    
    # plot the correlation matrix as a heatmap
    gg_out = ggplot(data = melt(curr_df)) + geom_tile(aes(x=variable,y=compare_samp, fill = value)) +
                    geom_text(aes(x=variable,y=compare_samp, label=round(value, 2))) + 
                    scale_fill_gradient2(low="blue", mid="white", high="red", 
                                        midpoint=0,    
                                        breaks=seq(-1,1,0.1), 
                                        limits=c(-1, 1)) + 
                    theme_bw() + labs(x = "NMF", y="K-Means") +
                    theme(legend.position = "none", 
                            axis.text.x=element_blank(),
                            axis.ticks.x=element_blank(),
                            axis.text.y=element_blank(),
                            axis.ticks.y=element_blank(),
                            axis.title = element_text(size = 15),
                            plot.title = element_text(size = 20, hjust = 0.5))

    return(gg_out)
}


plot_nmf_v_kmeans <- function(fig2_kmeans_files, fig2_nmf_files, dataset_id, dataset_idx, curr_k){
    
    curr_kmeans_file = fig2_kmeans_files[dataset_idx]
    curr_nmf_file = fig2_nmf_files[dataset_idx]
    
    kmeans_df = data.frame(fread(curr_kmeans_file))
    row.names(kmeans_df) = kmeans_df$V1
    kmeans_df = kmeans_df[,-1]
    
    nmf_df = data.frame(fread(curr_nmf_file))
    row.names(nmf_df) = nmf_df$V1
    nmf_df = nmf_df[,-1]
    
    stopifnot(all(colnames(kmeans_df) == colnames(nmf_df)))
    
    colnames(nmf_df) = paste(colnames(nmf_df), "NMF", sep="_")
    cor_df = data.frame(cor(kmeans_df, nmf_df))
    
    dataset_plot = plot_nmf_v_kmeans_internal(cor_df, curr_k)
    return(dataset_plot)
    
}


```

## Now plot NMF vs. Kmeans for each dataset

### AACES RNA-Seq

```{r aaces_kmeans_nmf, fig.align='center', fig.width=75, fig.height=15, echo=TRUE, warning=FALSE, message=F}

dataset_id = "aaces"
dataset_idx = 2
cluster_compare = c("ClusterK3")

aaces_plot = plot_nmf_v_kmeans(fig2_kmeans_files, fig2_nmf_files, dataset_id, dataset_idx, curr_k = "ClusterK3")
aaces_plot = aaces_plot + ggtitle("Schildkraut B")
aaces_plot

outfile = paste0(proj_dir, "/figure_notebooks/manuscript_figs/schildkrautAA_K3_kmeans_v_nmf.pdf")
ggsave(outfile,
       aaces_plot, width = 3, height = 3, units = "in", device = "pdf")


```

```{r aaces_kmeans_nmf_k3, fig.align='center', fig.width=75, fig.height=15, echo=TRUE, warning=FALSE, message=F}

dataset_id = "aaces"
dataset_idx = 3
cluster_compare = c("ClusterK3")

aaces_plot = plot_nmf_v_kmeans(fig2_kmeans_files, fig2_nmf_files, dataset_id, dataset_idx, curr_k = "ClusterK3")
aaces_plot = aaces_plot + ggtitle("Schildkraut W")
aaces_plot

outfile = paste0(proj_dir, "/figure_notebooks/manuscript_figs/schildkrautW_K3_kmeans_v_nmf.pdf")
ggsave(outfile,
       aaces_plot, width = 3, height = 3, units = "in", device = "pdf")


```



### GSE32062.GPL6480 RNA-Seq

```{r GSE32062.GPL6480_kmeans_nmf, fig.align='center', fig.width=75, fig.height=15, echo=TRUE, warning=FALSE, message=F}

dataset_id = "GSE32062.GPL6480"
dataset_idx = 5

dataset_plot = plot_nmf_v_kmeans(fig2_kmeans_files, fig2_nmf_files, dataset_id, dataset_idx, curr_k = "ClusterK3")
dataset_plot = dataset_plot + ggtitle("Yoshihara")
dataset_plot


outfile = paste0(proj_dir, "/figure_notebooks/manuscript_figs/yoshihara_K3_kmeans_v_nmf.pdf")
ggsave(outfile,
       dataset_plot, width = 3, height = 3, units = "in", device = "pdf")


```


### GSE9891 RNA-Seq

```{r GSE9891_kmeans_nmf, fig.align='center', fig.width=75, fig.height=15, echo=TRUE, warning=FALSE, message=F}

dataset_id = "GSE9891"
dataset_idx = 6

dataset_plot = plot_nmf_v_kmeans(fig2_kmeans_files, fig2_nmf_files, dataset_id, dataset_idx, curr_k = "ClusterK3")
dataset_plot = dataset_plot + ggtitle("Tothill")
dataset_plot


outfile = paste0(proj_dir, "/figure_notebooks/manuscript_figs/tothill_K3_kmeans_v_nmf.pdf")
ggsave(outfile,
       dataset_plot, width = 3, height = 3, units = "in", device = "pdf")

```


### mayo RNA-Seq

```{r mayo_kmeans_nmf, fig.align='center', fig.width=75, fig.height=15, echo=TRUE, warning=FALSE, message=F}

dataset_id = "mayo"
dataset_idx = 7

dataset_plot = plot_nmf_v_kmeans(fig2_kmeans_files, fig2_nmf_files, dataset_id, dataset_idx, curr_k = "ClusterK3")
dataset_plot = dataset_plot + ggtitle("Mayo")
dataset_plot


outfile = paste0(proj_dir, "/figure_notebooks/manuscript_figs/mayo_K3_kmeans_v_nmf.pdf")
ggsave(outfile,
       dataset_plot, width = 3, height = 3, units = "in", device = "pdf")


```


### TCGA RNA-Seq

```{r tcga_kmeans_nmf, fig.align='center', fig.width=75, fig.height=15, echo=TRUE, warning=FALSE, message=F}

dataset_id = "TCGA"
dataset_idx = 8

dataset_plot = plot_nmf_v_kmeans(fig2_kmeans_files, fig2_nmf_files, dataset_id, dataset_idx, curr_k = "ClusterK3")
dataset_plot = dataset_plot + ggtitle("TCGA")
dataset_plot


outfile = paste0(proj_dir, "/figure_notebooks/manuscript_figs/tcga_K3_kmeans_v_nmf.pdf")
ggsave(outfile,
       dataset_plot, width = 3, height = 3, units = "in", device = "pdf")

```


