---
title: "sample_heatmap_annotated"
author: "Natalie Davidson"
date: "9/12/2021"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

require("data.table")
require("here")
require("ggplot2")
require("dplyr")
require("SummarizedExperiment")
require("ggrepel")
require("RColorBrewer")
require("readr")
require("pheatmap")

proj_dir = here()
source(file.path(proj_dir, "/analyze_rnaseq/plot_utils.R"))
source(file.path(proj_dir, "/comparison/utils/file_processing_utils.R"))
set.seed(5)


```

## Sample Heatmap 

This notebook is for generating a heatmap comparing how correlated each samples is to another using log10(normalized counts)
These counts are the same values used in the Way pipeline.
The samples are then annotated by their cluster ID.
Cluster IDs from AACES are mapped to predefined TCGA clusters.

Cluster IDs are from: `/data/way_pipeline_results_10removed_NeoRemoved/2.Clustering_DiffExprs-Tables-ClusterMembership/FullClusterMembership.csv`

Cluster ID to TCGA subtype are from: `/data/way_pipeline_results_10removed_NeoRemoved/2.Clustering_DiffExprs-Tables-Data_Concordance/TCGA_KmeansClusterK*.csv

For the AACES samples, both of those tables are summarized and in the following metadata file:
Sample Metadata: `/reference_data/main_metadata_table.tsv`

Sample Expression: `/data/rna_seq_pilot_and_new/salmon_normalized_filtered_for_way_pipeline_bottom10Removed.tsv.gz`
Genes Used in clustering: `/data/way_pipeline_results_10removed_NeoRemoved/1.DataInclusion-Data-Genes/CommonGenes_genelist.csv`

### read all data

```{r read_data, fig.align='center', echo=TRUE, warning=FALSE, message=F}

# read in the expression
rnaseq_expr_file = file.path(proj_dir, 
                    "/data/rna_seq_pilot_and_new/salmon_normalized_filtered_for_way_pipeline_bottom10Removed.tsv.gz")
rnaseq_expr_df = data.frame(fread(rnaseq_expr_file))


# filter for genes used in the analysis
colnames(rnaseq_expr_df)[1] = "hgnc_symbol"
gene_file = file.path(proj_dir, 
                    "/data/way_pipeline_results_10removed_NeoRemoved/1.DataInclusion-Data-Genes/CommonGenes_genelist.csv")
gene_df <- data.frame(fread(gene_file))

colnames(gene_df)[1] = "hgnc_symbol"

rnaseq_expr_df = subset(rnaseq_expr_df, hgnc_symbol %in% gene_df$hgnc_symbol)
gene_ids = rnaseq_expr_df$hgnc_symbol

# format into a style for correlations
rnaseq_expr_df = rnaseq_expr_df[,-1]
sample_ids = colnames(rnaseq_expr_df)
rnaseq_expr_df = data.frame(t(rnaseq_expr_df))

# now add column names back so we can have annotations
colnames(rnaseq_expr_df) = gene_ids
rnaseq_expr_df$ID = gsub("Sample_", "", row.names(rnaseq_expr_df))

# read in the metadata
metadata_file = file.path(proj_dir, 
                    "/reference_data/main_metadata_table.tsv")
metadata_df = data.frame(fread(metadata_file))
full_df = merge(metadata_df, rnaseq_expr_df, by = "ID")


```


### make method to calculate correlations

```{r calc_corr_method, fig.align='center', echo=TRUE, warning=FALSE, message=F}

library(amap)

calc_corr <- function(full_df, k_idx, kmeans_colname, gene_ids, proj_dir){
    
    curr_df = full_df
    colname_idx = which(colnames(curr_df) == kmeans_colname)
    colnames(curr_df)[colname_idx] = "KMeans_Result"
    curr_df = curr_df[order(curr_df$KMeans_Result),]

    
    expr_df = subset(curr_df, KMeans_Result == 1)
    dist_mat1 <- Dist(log10(expr_df[,gene_ids]+1), method = 'pearson')
    hclust_1 <- hclust(dist_mat1, method = 'complete')$order
    
    expr_df = subset(curr_df, KMeans_Result == 2)
    dist_mat2 <- Dist(log10(expr_df[,gene_ids]+1), method = 'pearson')
    hclust_2 <- hclust(dist_mat2, method = 'complete')$order

    temp_df = rbind(subset(curr_df, KMeans_Result == 1)[hclust_1,],
                subset(curr_df, KMeans_Result == 2)[hclust_2,])

    if(k_idx > 2){
        expr_df = subset(curr_df, KMeans_Result == 3)
        dist_mat3 <- Dist(log10(expr_df[,gene_ids]+1), method = 'pearson')
        hclust_3 <- hclust(dist_mat3, method = 'complete')$order
        temp_df = rbind(temp_df,
                        subset(curr_df, KMeans_Result == 3)[hclust_3,])
    }
     if(k_idx > 3){
        expr_df = subset(curr_df, KMeans_Result == 4)
        dist_mat4 <- Dist(log10(expr_df[,gene_ids]+1), method = 'pearson')
        hclust_4 <- hclust(dist_mat4, method = 'complete')$order
        temp_df = rbind(temp_df,
                        subset(curr_df, KMeans_Result == 4)[hclust_4,])
    }
    curr_df = temp_df
    
    annotation_row_df <- data.frame(subset(curr_df, select=c(KMeans_Result)))
    annotation_row_df$KMeans_Result = paste0("Clust", annotation_row_df$KMeans_Result)
    annotation_row_df$KMeans_Result = as.factor(annotation_row_df$KMeans_Result)
    row.names(annotation_row_df) = curr_df$ID
    
    # set up colors
    cluster_colors = brewer.pal(n = 8, name = "Set2")
    if(k_idx == 2){
        ann_colors = list(
        KMeans_Result = c("Clust1"=cluster_colors[1],"Clust2"=cluster_colors[2]))
    }else if(k_idx == 3){
        ann_colors = list(
        KMeans_Result = c("Clust1"=cluster_colors[1], 
                            "Clust2"=cluster_colors[2],
                            "Clust3"=cluster_colors[3]))
    }else if(k_idx == 4){
        ann_colors = list(
        KMeans_Result = c("Clust1"=cluster_colors[1], 
                            "Clust2"=cluster_colors[2],
                            "Clust3"=cluster_colors[3],
                            "Clust4"=cluster_colors[4]))
    }else{
        print("k_idx invalid")
        return(NA)
    }
    
    # make the correlation matrix
    expr_df = curr_df[,gene_ids]
    expr_df = (t(log10(expr_df+1)))
    cor_df <- cor(expr_df, method="pearson", use="complete.obs")
    row.names(cor_df) = curr_df$ID
    colnames(cor_df) = curr_df$ID
    
    # now plot
    gg = pheatmap(cor_df, annotation_row = annotation_row_df, 
             annotation_col = annotation_row_df,
             annotation_colors = ann_colors,
             scale = "none", show_rownames = F,
             annotation_names_row = F,
             show_colnames = F, cluster_cols = F, cluster_rows = F,
             main = paste("K =", k_idx))
    
    outfile = paste0(proj_dir, "/figure_notebooks/manuscript_figs/", k_idx, "_full_heatmap.pdf")
    ggsave(outfile,
           gg, width = 7, height = 5, units = "in", device = "pdf")
    
    return(gg)
}

```

### calculate correlations for K = 2, 3, 4

```{r calc_corr, fig.align='center', echo=TRUE, warning=FALSE, message=F}


kmeans_2 = calc_corr(full_df, k_idx=2, kmeans_colname="ClusterK2_kmeans", gene_ids, proj_dir)
kmeans_3 = calc_corr(full_df, k_idx=3, kmeans_colname="ClusterK3_kmeans", gene_ids, proj_dir)
kmeans_4 = calc_corr(full_df, k_idx=4, kmeans_colname="ClusterK4_kmeans", gene_ids, proj_dir)


```


### format figure 

```{r make_fig1, echo=TRUE, warning=FALSE, message=F}


k2_gg = image_read_pdf(file.path(proj_dir,
                                  "/figure_notebooks/manuscript_figs/2_full_heatmap.pdf"))
k3_gg = image_read_pdf(file.path(proj_dir,
                                  "/figure_notebooks/manuscript_figs/3_full_heatmap.pdf"))
k4_gg = image_read_pdf(file.path(proj_dir,
                                  "/figure_notebooks/manuscript_figs/4_full_heatmap.pdf"))
k2_gg = image_annotate(k2_gg, "a", size = 160)
k3_gg = image_annotate(k3_gg, "b", size = 160)
k4_gg = image_annotate(k4_gg, "c", size = 160)

full_panel <- image_append(image_scale(c(k2_gg, k3_gg, k4_gg), 3000), stack = FALSE)

print(full_image)
outfile = file.path(proj_dir,"/figure_notebooks/manuscript_figs/full_heatmap.pdf")
image_write(full_image, format = "pdf", outfile)
outfile = file.path(proj_dir,"/figure_notebooks/manuscript_figs/full_heatmap.png")
image_write(full_image, format = "png", outfile)

```
