---
title: "check_affy_degradation"
author: "Natalie Davidson"
date: "9/21/2022"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

require("data.table")
require("here")
require("ggplot2")
require("dplyr")
require("ggpubr")
require("VennDiagram")
require("RColorBrewer")
require("Rtsne")
require("viridis")

library("curatedOvarianData")

theme_set(theme_pubr())

proj_dir = here()
source(file.path(proj_dir, "/analyze_rnaseq/plot_utils.R"))

set.seed(5)


```

In this notebook, we will show how the Tothill and TCGA datasets have clusters that are correlated with library size, but not fully explained by degradation.
Since these are the only 2 Affy datasets in our analysis, we are only able to calculate a degradtaion score for these datasets.
This is because Affy MAs (microarrays) use multiple probes per gene, where we can estimate 5'/3' bias.
The other datasets are done on agilent -- we can not calculate probe bias on those MA's, so we instead look at biases of gene length.


## read in metadata

```{r read_meta, fig.align='center', fig.width=15, fig.height=15, echo=TRUE, warning=FALSE, message=F}

# current cluster labels
clust_id_file = file.path(proj_dir, 
                    "/data/way_pipeline_results_all/2.Clustering_DiffExprs-Tables-ClusterMembership/FullClusterMembership.csv")

clust_df = data.frame(fread(clust_id_file))
colnames(clust_df)[1] = "ID"

# original cluster labels
old_file = file.path(proj_dir, 
                    "/data/way_publication_results/033514_tables3.csv")

old_df = data.frame(fread(old_file))

# gene length
gene_meta_file = file.path(proj_dir, 
                    "/reference_data/temp_gene_lengths.tsv")
gene_len_df = data.frame(fread(gene_meta_file))
gene_len_df = gene_len_df[order(gene_len_df$Median, decreasing = T),]
gene_len_df = gene_len_df[!duplicated(gene_len_df$HGNC.symbol),]

# MAD genes used in clustering
gene_file = file.path(proj_dir, 
                    "/data/way_pipeline_results_10removed_NeoRemoved/1.DataInclusion-Data-Genes/GlobalMAD_genelist.csv")
gene_df <- data.frame(fread(gene_file))
colnames(gene_df)[1] = "hgnc_symbol"


```




## Make boxplots methods

```{r boxplots_method, fig.align='center', fig.width=15, fig.height=15, echo=TRUE, warning=FALSE, message=F}



read_format_expr <- function(in_df, metadata_table){
    
     rnaseq_expr_df = in_df

    # format it so we can add metadata
    gene_ids = row.names(rnaseq_expr_df)
    sample_ids = colnames(rnaseq_expr_df)
    rnaseq_expr_df = data.frame(t(rnaseq_expr_df))
    colnames(rnaseq_expr_df) = gene_ids
    
    # now add column names back so we can have annotations
    rnaseq_expr_df$ID = sample_ids
    
    full_df = merge(metadata_table, rnaseq_expr_df, by = "ID")
    
    return(list(full_df, gene_ids))
}

plot_boxplot_expr <- function(in_df, in_genes, metadata_table, gene_len_df, col_name_interest, 
                              col_plot_name, title_txt, xlab_txt, ylab_txt){
        
    # now plot the stats
    gene_count_df = in_df[,in_genes]
    gene_count_df$sample_id = in_df$ID
    gene_count_melt <- melt(gene_count_df, id.vars="sample_id")
    colnames(gene_count_melt) = c("ID", "gene_id", "count")
    
    # add version tag
    gene_count_melt = merge(gene_count_melt, metadata_table, by="ID")
    
    # order samples by median expression
    sample_order_df = data.frame(median_count = apply(in_df[,in_genes], 1, median),
                                 ID = in_df$ID)
    sample_order_df = sample_order_df[order(sample_order_df$median_count),]
    gene_count_melt = merge(gene_count_melt, sample_order_df)
    gene_count_melt$ID = factor(gene_count_melt$ID, 
                                        levels = sample_order_df$ID)

    col_idx = which(colnames(gene_count_melt) == col_name_interest)
    colnames(gene_count_melt)[col_idx] = "fill_col"
    gg = ggplot(gene_count_melt, aes(x=as.factor(ID), y=count, fill=as.factor(fill_col))) +
        geom_boxplot() + theme_bw() +
        theme(axis.text.x = element_blank()) +
        xlab(xlab_txt) + ylab(ylab_txt) +
            ggtitle(title_txt) +
        labs(fill = col_plot_name)
    ggS = ggplot(gene_count_melt, aes(x=as.factor(ID), y=count, fill=round(slope, 1))) +
        geom_boxplot() + theme_bw() +
        theme(axis.text.x = element_blank()) +
        xlab(xlab_txt) + ylab(ylab_txt) +
        ggtitle(title_txt) + scale_fill_viridis() +
        labs(fill = "RNADeg (lower\n= degraded)")   
    
    med_df = unique(gene_count_melt[,c("ID", "fill_col", "median_count", "slope")])
    gg1 = ggplot(med_df, aes(x=fill_col, y=median_count, fill=as.factor(fill_col))) +
            geom_boxplot() + theme_bw() +
            xlab(col_plot_name) + ylab("median") +
            labs(fill = col_plot_name) + theme(legend.position = "none", axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))


    # subset for the short genes only
    gene_len_df_inter = intersect(gene_count_melt$gene_id, gene_len_df$HGNC.symbol)
    gene_len_df_inter = subset(gene_len_df, HGNC.symbol %in% gene_len_df_inter)
    cutoff = quantile(gene_len_df_inter$Median, 0.10)
    gene_len_df_short = subset(gene_len_df_inter, Median < cutoff)
    cutoff = quantile(gene_len_df_inter$Median, 0.90)
    gene_len_df_long = subset(gene_len_df, Median > cutoff)

    gene_count_melt_short = subset(gene_count_melt, gene_id %in% gene_len_df_short$HGNC.symbol)
    gene_count_melt_short_med = gene_count_melt_short %>%
                                    group_by(ID)%>% 
                                    summarise(Median_short=median(count))
    gene_count_melt_short = merge(gene_count_melt_short, gene_count_melt_short_med)
    x_lab_gene = paste(col_plot_name, "Ngenes=", length(unique(gene_count_melt_short$gene_id)))
    gg2 = ggplot(gene_count_melt_short, aes(x=fill_col, y=Median_short, fill=as.factor(fill_col))) +
            geom_boxplot() + theme_bw() +
            xlab(x_lab_gene) + ylab("short genes median ") +
            labs(fill = col_plot_name) + theme(legend.position = "none", axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
    
    
    gene_count_melt_long = subset(gene_count_melt, gene_id %in% gene_len_df_long$HGNC.symbol)
    gene_count_melt_long_med = gene_count_melt_long %>%
                                    group_by(ID)%>% 
                                    summarise(Median_long=median(count))
    gene_count_melt_long = merge(gene_count_melt_long, gene_count_melt_long_med)
    x_lab_gene = paste(col_plot_name, "Ngenes=", length(unique(gene_count_melt_long$gene_id)))
    gg3 = ggplot(gene_count_melt_long, aes(x=fill_col, y=Median_long, fill=as.factor(fill_col))) +
            geom_boxplot() + theme_bw() +
            xlab(x_lab_gene) + ylab("long genes median ") +
            labs(fill = col_plot_name) + theme(legend.position = "none", axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

    gg4 = ggplot(med_df, aes(slope, fill=as.factor(fill_col))) +
            geom_density(alpha=0.5) + theme_bw() +
            xlab(col_plot_name) + ylab("RNADeg (lower\n= degraded)") +
            labs(fill = col_plot_name) + theme(legend.position = "none")
    gg5 = ggplot(med_df, aes(x=fill_col, y=slope, fill=as.factor(fill_col))) +
            geom_boxplot() + theme_bw() +
            xlab(col_plot_name) + ylab("RNADeg (lower\n= degraded)") +
            labs(fill = col_plot_name) + theme(legend.position = "none", axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

    flat_plot = ggarrange(gg, ggS, 
                          ggarrange(gg1, gg2, gg3, gg4, gg5, ncol=5),
                          nrow=3)
    return(flat_plot)
}


```

# Tothill


## read in tothill

```{r read_tothill, fig.align='center', fig.width=15, fig.height=15, echo=TRUE, warning=FALSE, message=F}

tothill_data = data(GSE9891_eset)
ExpressionData <- get(tothill_data)
tothill_dta <- exprs(ExpressionData)

tothill_metadata_table = subset(clust_df, Dataset == "Tothill")


tothill_old_df = subset(old_df, Dataset == "Tothill")
tothill_old_df = tothill_old_df[,-2]
colnames(tothill_old_df)[2:ncol(tothill_old_df)] = paste0("ORIG_", colnames(tothill_old_df)[2:ncol(tothill_old_df)])
tothill_metadata_table = merge(tothill_metadata_table, tothill_old_df, by="ID")

# read in the degradation scores
slope_file = file.path(proj_dir, 
                    "/reference_data/tothill_rnadeg.tsv")
slope_df = data.frame(fread(slope_file))
colnames(slope_df) = c("ID", "rna_deg")
slope_df$ID = gsub(".CEL.gz", "", slope_df$ID)
tothill_metadata_table = merge(tothill_metadata_table, slope_df)

# read in the other method for calculating degradation scores
slope_file = file.path(proj_dir, 
                    "/reference_data/tothill_slope.tsv")
slope_df = data.frame(fread(slope_file))
colnames(slope_df) = c("ID", "slope", "threeP", "fiveP")
tothill_metadata_table = merge(tothill_metadata_table, slope_df)



```



## Make boxplots 

```{r boxplots, fig.align='center', fig.width=15, fig.height=15, echo=TRUE, warning=FALSE, message=F}


# format the expr table
res = read_format_expr(tothill_dta, tothill_metadata_table)
in_df = res[[1]]
in_genes = res[[2]]


subset_cols = c(colnames(tothill_metadata_table), gene_df$hgnc_symbol)


raw_gg = plot_boxplot_expr(in_df[,subset_cols], gene_df$hgnc_symbol, tothill_metadata_table, gene_len_df,
                  col_name_interest = "ClusterK3_kmeans", 
                  col_plot_name = "K3_kmeans_clusters",
                  title = "tothill Intensity distribution for each sample", 
                  xlab = "Samples", 
                  ylab = "Intensity")
    
outfile = paste0(proj_dir, "/figure_notebooks/manuscript_figs/K3_tothill_degradation.pdf")
ggsave(outfile,
       raw_gg, width = 30, height = 15, units = "in", device = "pdf")
raw_gg

### K=2

raw_gg = plot_boxplot_expr(in_df[,subset_cols], gene_df$hgnc_symbol, tothill_metadata_table, gene_len_df,
                  col_name_interest = "ClusterK2_kmeans", 
                  col_plot_name = "K2_kmeans_clusters",
                  title = "tothill Intensity distribution for each sample", 
                  xlab = "Samples", 
                  ylab = "Intensity")
    
outfile = paste0(proj_dir, "/figure_notebooks/manuscript_figs/K2_tothill_degradation.pdf")
ggsave(outfile,
       raw_gg, width = 30, height = 15, units = "in", device = "pdf")
raw_gg


####### calculate test

sample_order_df = data.frame(med = apply(in_df[,in_genes], 1, median),
                             ID = in_df$ID)

wilcox_df = merge(sample_order_df, tothill_metadata_table, by="ID")
wilcox_df = wilcox_df[order(wilcox_df$slope),]

num_g1 = sum(wilcox_df$ClusterK2_kmeans == 1)

print(wilcox.test(wilcox_df$ClusterK2_kmeans[1:num_g1],
                  wilcox_df$ClusterK2_kmeans[(num_g1+1):nrow(wilcox_df)]))
print(wilcox.test(wilcox_df$ClusterK2_NMF[1:num_g1], 
                  wilcox_df$ClusterK2_NMF[(num_g1+1):nrow(wilcox_df)]))

    
    
```


# TCGA


## read in TCGA

```{r read_tcga, fig.align='center', fig.width=15, fig.height=15, echo=TRUE, warning=FALSE, message=F}

tcga_data = data(TCGA_eset)
ExpressionData <- get(tcga_data)
tcga_dta <- exprs(ExpressionData)

tcga_metadata_table = subset(clust_df, Dataset == "TCGA")


tcga_old_df = subset(old_df, Dataset == "TCGA")
tcga_old_df = tcga_old_df[,-2]
colnames(tcga_old_df)[2:ncol(tcga_old_df)] = paste0("ORIG_", colnames(tcga_old_df)[2:ncol(tcga_old_df)])
tcga_metadata_table = merge(tcga_metadata_table, tcga_old_df, by="ID")


# get TCGA publication IDs
tcgaPub_file = file.path(proj_dir, 
                    "/reference_data/Verhaak2013_Supplemental.csv")

tcgaPub_df = data.frame(fread(tcgaPub_file))
tcgaPub_df = tcgaPub_df[-1,]
colnames(tcgaPub_df) = tcgaPub_df[1,]
tcgaPub_df = tcgaPub_df[-1,]
tcgaPub_df = subset(tcgaPub_df, DATASET %in% c("TCGA-discovery", "TCGA-validation"))
tcgaPub_df$Dataset = "TCGA"
tcgaPub_df = tcgaPub_df[,c("ID", "SUBTYPE")]
tcgaPub_df$ID = gsub("-", ".", tcgaPub_df$ID)

tcga_metadata_table = merge(tcga_metadata_table, tcgaPub_df, by="ID")

# get the slope of 3'/5'
slope_file = file.path(proj_dir, 
                    "/reference_data/tcga_slope.tsv")
slope_df = data.frame(fread(slope_file))
slope_df_dup = which(duplicated(slope_df$ID))
slope_df_dup = slope_df$ID[slope_df_dup]
slope_df = subset(slope_df, ! ID %in% slope_df_dup )
colnames(slope_df)[2] = c("slope")
tcga_metadata_table = merge(tcga_metadata_table, slope_df, by="ID")

# get the rnadeg
slope_file = file.path(proj_dir, 
                    "/reference_data/tcga_rnadeg.tsv")
slope_df = data.frame(fread(slope_file))

#format rna_deg
colnames(slope_df) = c("cel_ID", "rna_deg")
slope_df$cel_ID = gsub(".CEL", "", slope_df$cel_ID)

# reformat the IDs in the degradation file
tcga_samp_file = file.path(proj_dir, 
                    "/reference_data/broad.mit.edu_OV.HT_HG-U133A.sdrf.txt")
tcga_samp_df = data.frame(fread(tcga_samp_file))
tcga_samp_df = unique(tcga_samp_df[,c("Comment..TCGA.Barcode.", "Hybridization.Name")])
colnames(tcga_samp_df) = c("ID", "cel_ID")
tcga_samp_df$ID = substr(tcga_samp_df$ID, 1, 12)
tcga_samp_df$ID = gsub("-", ".", tcga_samp_df$ID)
slope_df = merge(slope_df, tcga_samp_df, by="cel_ID")
slope_df = slope_df[,c("rna_deg", "ID")]

# remove duplicated aliquots -- this is too restrictive
# but this is just for sanity checks 
# if we want to publish this analysis we will have to do a better job matching
# sample IDs to aliquots
slope_df_dup = which(duplicated(slope_df$ID))
slope_df_dup = slope_df$ID[slope_df_dup]
slope_df = subset(slope_df, ! ID %in% slope_df_dup )
tcga_metadata_table = merge(tcga_metadata_table, slope_df, by="ID")



```



## Make boxplots 

```{r boxplots_tcga, fig.align='center', fig.width=15, fig.height=15, echo=TRUE, warning=FALSE, message=F}


# format the expr table
res = read_format_expr(tcga_dta, tcga_metadata_table)
in_df = res[[1]]
in_genes = res[[2]]


subset_cols = c(colnames(tcga_metadata_table), gene_df$hgnc_symbol)


raw_gg = plot_boxplot_expr(in_df[,subset_cols], gene_df$hgnc_symbol, tcga_metadata_table, gene_len_df,
                  col_name_interest = "ClusterK3_kmeans", 
                  col_plot_name = "K3_kmeans_clusters",
                  title = "tcga Intensity distribution for each sample", 
                  xlab = "Samples", 
                  ylab = "Intensity")
    
outfile = paste0(proj_dir, "/figure_notebooks/manuscript_figs/K3_tcga_degradation.pdf")
ggsave(outfile,
       raw_gg, width = 30, height = 15, units = "in", device = "pdf")
raw_gg

### K=2

raw_gg = plot_boxplot_expr(in_df[,subset_cols], gene_df$hgnc_symbol, tcga_metadata_table, gene_len_df,
                  col_name_interest = "ClusterK2_kmeans", 
                  col_plot_name = "K2_kmeans_clusters",
                  title = "tcga Intensity distribution for each sample", 
                  xlab = "Samples", 
                  ylab = "Intensity")
    
outfile = paste0(proj_dir, "/figure_notebooks/manuscript_figs/K2_tcga_degradation.pdf")
ggsave(outfile,
       raw_gg, width = 30, height = 15, units = "in", device = "pdf")
raw_gg

# plot the original TCGA paper clusters
raw_gg = plot_boxplot_expr(in_df[,subset_cols], gene_df$hgnc_symbol, tcga_metadata_table, gene_len_df,
                  col_name_interest = "SUBTYPE", 
                  col_plot_name = "SUBTYPE",
                  title = "TCGA Intensity distribution for each sample", 
                  xlab = "Samples", 
                  ylab = "Intensity")
    
outfile = paste0(proj_dir, "/figure_notebooks/manuscript_figs/TCGAPUB_TCGA.pdf")
ggsave(outfile,
       raw_gg, width = 30, height = 15, units = "in", device = "pdf")
raw_gg


####### calculate test

sample_order_df = data.frame(med = apply(in_df[,in_genes], 1, median),
                             ID = in_df$ID)

wilcox_df = merge(sample_order_df, tcga_metadata_table, by="ID")
wilcox_df = wilcox_df[order(wilcox_df$slope),]

num_g1 = sum(wilcox_df$ClusterK2_kmeans == 1)

print(wilcox.test(wilcox_df$ClusterK2_kmeans[1:num_g1],
                  wilcox_df$ClusterK2_kmeans[(num_g1+1):nrow(wilcox_df)]))
print(wilcox.test(wilcox_df$ClusterK2_NMF[1:num_g1], 
                  wilcox_df$ClusterK2_NMF[(num_g1+1):nrow(wilcox_df)]))

    
    
```


# Bonome


## read in Bonome

```{r read_bonome, fig.align='center', fig.width=15, fig.height=15, echo=TRUE, warning=FALSE, message=F}

bonome_data = data(GSE26712_eset)
ExpressionData <- get(bonome_data)
bonome_dta <- exprs(ExpressionData)

bonome_metadata_table = subset(old_df, Dataset == "Bonome")

# read in the other method for calculating degradation scores
slope_file = file.path(proj_dir, 
                    "/reference_data/bonome_slope.tsv")
slope_df = data.frame(fread(slope_file))
colnames(slope_df) = c("ID", "slope", "threeP", "fiveP")
bonome_metadata_table = merge(bonome_metadata_table, slope_df)



```


## Make boxplots 

```{r boxplots_bonome, fig.align='center', fig.width=15, fig.height=15, echo=TRUE, warning=FALSE, message=F}


# format the expr table
res = read_format_expr(bonome_dta, bonome_metadata_table)
in_df = res[[1]]
in_genes = res[[2]]


subset_cols = c(colnames(bonome_metadata_table), gene_df$hgnc_symbol)


raw_gg = plot_boxplot_expr(in_df[,subset_cols], gene_df$hgnc_symbol, bonome_metadata_table, gene_len_df,
                  col_name_interest = "ClusterK3_kmeans", 
                  col_plot_name = "K3_kmeans_clusters",
                  title = "Bonome Intensity distribution for each sample", 
                  xlab = "Samples", 
                  ylab = "Intensity")
    
outfile = paste0(proj_dir, "/figure_notebooks/manuscript_figs/K3_bonome_degradation.pdf")
ggsave(outfile,
       raw_gg, width = 30, height = 15, units = "in", device = "pdf")
raw_gg

### K=2

raw_gg = plot_boxplot_expr(in_df[,subset_cols], gene_df$hgnc_symbol, bonome_metadata_table, gene_len_df,
                  col_name_interest = "ClusterK2_kmeans", 
                  col_plot_name = "K2_kmeans_clusters",
                  title = "bonome Intensity distribution for each sample", 
                  xlab = "Samples", 
                  ylab = "Intensity")
    
outfile = paste0(proj_dir, "/figure_notebooks/manuscript_figs/K2_bonome_degradation.pdf")
ggsave(outfile,
       raw_gg, width = 30, height = 15, units = "in", device = "pdf")
raw_gg


####### calculate test

sample_order_df = data.frame(med = apply(in_df[,in_genes], 1, median),
                             ID = in_df$ID)

wilcox_df = merge(sample_order_df, bonome_metadata_table, by="ID")
wilcox_df = wilcox_df[order(wilcox_df$slope),]

num_g1 = sum(wilcox_df$ClusterK2_kmeans == 1)

print(wilcox.test(wilcox_df$ClusterK2_kmeans[1:num_g1],
                  wilcox_df$ClusterK2_kmeans[(num_g1+1):nrow(wilcox_df)]))
print(wilcox.test(wilcox_df$ClusterK2_NMF[1:num_g1], 
                  wilcox_df$ClusterK2_NMF[(num_g1+1):nrow(wilcox_df)]))

    
    
```
